---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nonogram

<!-- badges: start -->
`r badger::badge_devel("hrj21/nonogram", "blue")`
`r badger::badge_lifecycle("stable")`
<!-- badges: end -->

R package for solving black and white nonogram puzzles.

## Installation

You can install the nonogram package from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("hrj21/nonogram")
```

## Example of defining, solving, and plotting a simple 3x3 nonogram

The package defines an S3 class `nonogram` used for representing nonogram puzzles. The `nonogram()` constructor function allows you to define a nonogram by hand. The `rows` and `columns` arguments should be lists of numeric vectors representing the patterns in each row and column (starting with the top row and leftmost column).

```{r define_3x3, eval=FALSE}
library(nonogram)

non_3x3 <- nonogram(
  rows    = list(c(1, 1), 1, 3),
  columns = list(c(1, 1), 2, c(1, 1))
)

non_3x3

```

There are currently two algorithms implemented for solving nonograms: a brute force algorithm, and a permutation elimination algorithm. The brute force algorithm calculates the porportion of filled squares in the nonogram, and keeps proposing random matrices of 0s and 1s until the row and column patterns of a proposal match those of the nonogram in question.

To solve our nono_3x3 nonogram using the brute force algorithm, we simply call the `solve()` method, specifying `algorithm = "force"`. You will need to set the `max_iter` argument higher than the default (10). 

```{r solve_3x3, eval=FALSE}
non_3x3 <- solve(non_3x3, algorithm = "force", max_iter = Inf)
# ?solve.nonogram
```

You can plot the solved nongram by passing it to the `plot()` method.

You can also embed plots, for example:

```{r plot_3x3, eval=FALSE}
plot(non_3x3)
# ?plot.nonogram
```

## Example of solving bigger nonograms

As the number of possible solutions is 2^(n x m), the brute force algorithm is rarely useful beyond very small nonograms. Technically the number of solutions is smaller than 2^(n x m) when we constrain the proposal matrix to have the same proportion of 1s as the nonogram to be solved, but still the number of possible solutions becomes quickly intractable with even moderately-sized nonograms.

The other algorithm is the permutation elimination algorithm, which proceeds as follows:

    \enumerate{
        \item{For a nonogram with m rows and n columns, if m = n then generate 
            a single matrix with n columns whose rows represent every possible 
            permutation of 0s and 1s for a vector of length n. If m != n then 
            generate two matrices, one containing all possible row permutations
            and one containing all possible column permutations.}
        \item{Convert each permutation from the full permutation matrix/matrices
            into its run length-encoded form. E.g. 000110111101 would become
            241 in run length encoding.}
        \item{For each row and column of the nonogram, find every permutation in
            the full permutation matrix appropriate for that dimension, whose rows 
            have the same run length encoding as the nonogram's row/column in question.
            Set these permutations as the initial solution set for that row/column.
            Repeat this process until every row and column has an initial solution set.}
        \item{For each row and column of the nonogram, identify any elements across
            the solution set that are all 0 or all 1 for all possible permutations.
            E.g. if all permutations for column j have a 0 at element k, then eliminate 
            any solutions in row k that don't have a 0 at element j. Perform this 
            action for every row and column. A single iteration has passed once
            this has been applied to every row and column once.}
        \item{Repeat step 4 until only a single solution remains for every row and
            column. Return the nonogram with the solution.}
    }

Unless you have a specific nonogram you wish to solve or create, defining nonograms by hand can become a little tedious. Thankfully, the package comes with the `examples` dataset, which is a list of many predefined nonograms of various sizes. Let's pick one to solve.

```{r examples, eval=FALSE}
data(examples)

ex36 <- examples[[36]]
ex36
plot(ex36, with_solution = FALSE)

```

Let's solve our 15 x 15 nonogram using the permutation elimination algorithm (the default). In fact, all the arguments in the call to `solve` below are the defaults, except the nonogram itself.

```{r solve_ex36, eval=FALSE}
ex36 <- solve(ex36, algorithm = "perm_elim", verbose = TRUE, parallel = TRUE)
plot(ex36)
```
The main issue with this algorithm is that all possible permutations of 0s and 1s for the row and column lengths must be calculated up front. This quickly consumes a lot of memory and processing power and so this algorithm doesn't scale to puzzles larger than 30 x 30.

Even if solving smaller puzzles, if you are solving multiple puzzles of the same size, it is more efficient to calculate these full permutation matrices once, rather than separately for each puzzle. Let's say we wish to solve puzzles that are 15 x 15, we can calculate the matrix of all possible permutations of 0 and 1 of length 15 using the function `make_full_perm_set()`.

```{r full_perm, eval=FALSE}
perms15 <- make_full_perm_set(15)
perms15[1:10, ] # just the first 10 rows
dim(perms15)
```

```{r solve_pre_perms, eval = FALSE}
ex36 <- ex36 |> 
  solve(row_permutations = perms15, column_permutations = perms15) |> 
  plot()
```
## Conclusion

This package does not implement the best or most efficient nonogram solving algorithms, but you are very welcome to create a pull request to improve the existing package and implement better algorithms (such as those discussed at <https://webpbn.com/survey/>). This also isn't the only R package for solving nonograms. Please also check out <https://github.com/coolbutuseless/nonogram>.
